Supplementary material to the paper 'A simulation-based assessment of the relation between Stone Age sites and relative sea-level change along the Norwegian Skagerrak coast

### 2022-10-20 analysis/paper/paper.Rmd
Working on comments from peer reviewers

An exponential function has been fit to the distributions for vertical distance using maximum likelihood estimation (Figure \@ref(fig:results2)). While it makes theoretical sense that a process of exponential decay explains this relationship, it is also clear that this does not perfectly match the data. However, this can at least in part be related to methodological factors, where the accumulation of distance-values on the 0m mark likely follow from forcing negative values to zero, from the resolution of the spatial data, and from defining intersecting sea- and site polygon as having a zero distance. If one accepts this, the probability density function for exponential decay can be used to characterise the vertical distance between sites and the shoreline and be used to inform a method for shoreline dating that takes this into account.

>>>>

It is clear from the distributions that the simulation results have a severe right skew. Most sites were likely situated less than a meter from the shoreline, and from this there is a decline in density as one moves further along the x-axis. This trend is reasonably approximated by a exponential function that has been fit to the distributions for vertical distance using maximum likelihood estimation (Figure \@ref(fig:results2)). An expanded consideration of this relationship and comparison with some standard models is provided in the supplementary material. While exponential decay reasonably captures this relationship, it is also clear that it does not perfectly match the data. However, this can at least in part be related to methodological factors, where the accumulation of distance-values on the 0m mark likely follow from the resolution of the spatial data, from defining intersecting sea- and site polygon as having a zero distance, from sampling error and, although they are relatively few, from forcing negative values to zero. If one accepts this, the probability density function for exponential decay can be used to characterise the vertical distance between sites and the shoreline and be used to inform a method for shoreline dating that takes this into account.

### 2022-10-20 analysis/supplementary-materials/supplementary.Rmd
Working on comments from peer reviewers

Vallermyrene 1b is the younger than Vallermyrene 1a (above), as indicated both by the radiocarbon date, typology and elevation [@reitan2014f]. The DTM did not require editing.

@@

>>>>

Vallermyrene 1b is younger than Vallermyrene 1a (above), as indicated both by the radiocarbon date, typology and elevation [@reitan2014f]. The DTM did not require editing.

@@

## Model comparison

This section provides some further elaboration on the choice of an exponential model to describe the vertical distance between sites and shoreline. The normal distribution has been fit when including the negative values. After this, two exponential models were fit to the distrbutions. One when forcing negative values to zero and one where these are made NA, to evaluate the relative impact of these different choices. As the differences are negligeble, the choice was made to use the first solution. This is also the to which the power law distribution has been fit. To be able to fit a power law distribution, 0.01 have been added to all values, including for the normal and exponential functions, to avoid having values of zero. All models are fit using maximum likelihood estimation.

Although it appears evident that the exponential model is the best fit simply from inspecting the, the models have also been compared using the Schwarz criterion, often also termed the Bayesian information criteria [BIC, see e.g. @timpson2021; @crema2019 for previous use in archaeology]. This evaluates the ability of the models to describe the data while also taking into account the number of parameters in the models to avoid over-fitting.
